Waking a sleeping rabbit:\\ On Natural-Language Sentence Generation with FF

### 2009-12-18 trunk/Papers/crisp-ff-10/crisp-ff.tex
..

The goal agenda is completely off-target in our domain. The most
striking point concerns the interaction between the goal to create a
sentence ($\neg \mathsf{subst}(S,\mathsf{root})$ in the rabbit
example), and the goal to express a certain meaning
($\mathsf{sleep}(e,r_1)$ in the rabbit example). The goal agenda
orders the former before the latter because it detects that, once a
meaning is expressed, the sentence creation has started and it is not
possible to start anew anymore. Unless the inverse also
holds,\footnote{The inverse holds if, in the given grammar, the only
  way to create a meaning is to create a full sentence -- i.e., there
  are no side sentences like ``the sleeping rabbit''.}
$\mathsf{subst}(S,\mathsf{root})$ is alone in the first goal agenda
entry, i.e., there is no goal stating the actual meaning the sentence
is supposed to have. Consequently, the planner decides to generate
{\em any} sentence. If it happens to be the wrong sentence, the
planner is stuck in a dead-end. 
%While that dead end is recognized by
%the relaxed plan heuristic -- i.e., it is easy to see that the actions
%needed for achieving the goal cannot be applied anymore -- certainly
%this is an entirely unintended (and a bit amusing) outcome of this
%technique.
% Switching it off, we {\bf \dots
%example before/after.}

It is not clear to us, at this point, how goal ordering techniques
could be modified so that they return sensible results in the sentence
generation domain. The issue explained above regarding $\neg
\mathsf{subst}(S,\mathsf{root})$ and $\mathsf{sleep}(e,r_1)$ could be
addressed by recognizing that, once the latter is achieved, the former
is already true anyway. More generally however, the life-cycle of
substitution nodes and distractors is weird -- as we outline below, it
follows a false-true-false pattern -- and seems difficult to detect
based on domain analysis.

>>>>

The goal agenda is completely off-target in our domain. The most
striking point concerns the interaction between the goal to create a
sentence ($\neg \mathsf{subst}(S,\mathsf{root})$ in the rabbit
example), and the goal to express a certain meaning
($\mathsf{sleep}(e,r_1)$ in the rabbit example). The goal agenda
orders the former before the latter because it detects that, once a
meaning is expressed, the sentence creation has started and it is not
possible to start anew anymore. Unless the inverse also
holds,\footnote{The inverse holds if, in the given grammar, the only
  way to create a meaning is to create a full sentence -- i.e., there
  are no side sentences like ``the sleeping rabbit''.}
$\mathsf{subst}(S,\mathsf{root})$ is alone in the first goal agenda
entry, i.e., there is no goal stating the actual meaning the sentence
is supposed to have. Consequently, the planner decides to generate
{\em any} sentence. If it happens to be the wrong sentence, the
planner is stuck in a dead-end. %
%While that dead end is recognized by
%the relaxed plan heuristic -- i.e., it is easy to see that the actions
%needed for achieving the goal cannot be applied anymore -- certainly
%this is an entirely unintended (and a bit amusing) outcome of this
%technique.
% Switching it off, we {\bf \dots
%example before/after.}
It is not clear to us, at this point, how goal ordering techniques
could be modified so that they return more sensible results.

%The issue regarding $\neg
%\mathsf{subst}(S,\mathsf{root})$ and $\mathsf{sleep}(e,r_1)$ could
%potentially be addressed by recognizing that, once the latter is
%achieved, the former is already true anyway. Whether such techniques
%suffice to obtain useful goal orderings overall remains to be seen.

%More generally however,
%as outlined below the life-cycle of substitution nodes and distractors
%is weird; %
% -- as we
%outline below, it follows a false-true-false pattern --
%it seems difficult to detect this based on domain analysis.
